---
title: "Subclonal deconvolution of MSeq multi-region WGS colorectal cancer data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MSeq 2.0}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=T, message=FALSE, warning=FALSE, include=T}
library(evoverse)
library(evoverse.datasets)
```

This vignette describes the analysis of the MSeq data released with the paper 

* Caravagna et al. Model-based tumor subclonal reconstruction. biorxiv 2019, [doi:10.1101/586560](https://www.biorxiv.org/content/10.1101/586560v1).

The data is also released as Excel Supplementary Table with the paper.


### Available data

Mutation, copy nunmber ad purity estimates for both analysed multi-region samples - `Set6` and `Set7` - are part of the data released with the `evoverse.datasets` [package](https://github.com/caravagn/evoverse.datasets).



````{r}
# All available data
evoverse.datasets::available_data()

# Set7
data('MSEQ_CRC_ADENOCARCINOMA_SET7', package = 'evoverse.datasets')
```

Sneak peek to mutations, copy number, purity data and sample estimates

```{r}
print(MSEQ_CRC_ADENOCARCINOMA_SET7)
```

### Analysis

First create a dataset from the `evoverse`, retaining only diploid chromosomes.

```{r}
# Diploid segments 
segments_karyo_diploid = MSEQ_CRC_ADENOCARCINOMA_SET7$CNA %>% 
  select(-chr, -from, -to) %>% 
  apply(1, function(x) all(x == 1))

print(segments_karyo_diploid)

# Actual dataset
Set7 = evoverse::dataset(
  mutations = MSEQ_CRC_ADENOCARCINOMA_SET7$mutations, 
  segments = MSEQ_CRC_ADENOCARCINOMA_SET7$CNA[segments_karyo_diploid, ],
  samples = MSEQ_CRC_ADENOCARCINOMA_SET7$samples,
  purity = MSEQ_CRC_ADENOCARCINOMA_SET7$purity,
  description = "Set7 Sample from MSeq (multi-region WGS ~100x)"  
  )

print(Set7)
```

Basic information can be accessed through the other functions of the `evoverse` packages. For instance, you can visualise CNA data through
the  `CNAqc` [package](https://caravagn.github.op/CNAq).

```{r, fig.width=10, fig.height=8, warning=F}
# Example CNA rofiles with CNAqc
print(Set7$CNAqc$Set7_55)

# CNA segments plot
ggpubr::ggarrange(
  plotlist = lapply(Set7$CNAqc, function(x) { CNAqc::plot_segments(x) + ylim(0, 3) }),
  ncol = 1, 
  nrow = 4
)
```

We can print also the VAF of the annotated mutations, in both 2D and 1D type of plots using `evoverse`.

```{r, fig.width=9, fig.height=6, warning=F}
# plot by default is 2D
ggpubr::ggarrange(
  plotlist = plot(Set7),
  ncol = 3, 
  nrow = 2
)
```

``` {r, fig.width = 6, fig.height=6, warning=F}
# Using evoverse plot_1D_VAF (1D)
ggpubr::ggarrange(
  plotlist = lapply(Set7$samples, evoverse::plot_1D_VAF, x = Set7),
  ncol = 2, 
  nrow = 2
)
```

The following steps can be run to analyse the cohort (here, they are not run).
On a standard laptop this analysis takes about 10 minutes.

```{r,eval=FALSE}
# MOBSTER analysis with default parameters
Set7 =  evoverse::analyze_mobster(Set7)

# Save results
save(Set7, file = "MOBSTER_Set7_fits.RData")

# Plots of model selection that can be saved as PDF 
figure = lapply(Set7$fit_MOBSTER, mobster::plot_model_selection) 
  
pdf("MOBSTER_Set7_fits.pdf", width = 11, height = 9)
lapply(figure, print)
dev.off()

# Remove tail mutations and replot the data
Set7_notail = filter_tails(Set7)

plot(Set7_notail)

# VIBER clustering of Binomial mixtures for non-tail mutations
Set7_notail = evoverse::analyze_VIBER(Set7_notail, K = 7, alpha_0 = 1e-06, max_iter = 200)

save(Set7_notail, file = 'MOBSTER_VIBER_Set7_fits.RData')
  
# Plot 
pdf("MOBSTER_VIBER_Set7_fits_raw.pdf", width = 4, height = 3)
lapply(plot(Set7_notail, clusters = 'VIBER'), print)
dev.off()
  
# Apply heuristic (default parameters for instance) and replot
reduced_mv = VIBER::choose_clusters(Set7_notail$fit_VIBER)

pdf("MOBSTER_VIBER_Set7_fits_heuristics.pdf", width = 4, height = 3)
lapply(VIBER:::plot.vb_bmm(reduced_mv), print)
dev.off()
```

### Without MOBSTER

For comparison, run this.

```{r,eval=FALSE}
# VIBER clustering of Binomial mixtures for all mutations
Set7_standard = evoverse::analyze_VIBER(Set7, K = 9, alpha_0 = 1e-06, max_iter = 200)

save(Set7_standard, file = 'VIBER_Set7_fits.RData')
  
# Plot 
pdf("VIBER_Set7_fits_raw.pdf", width = 4, height = 3)
lapply(plot(Set7_standard, clusters = 'VIBER'), print)
dev.off()
  
# Apply heuristic (default parameters) and replot
reduced_mv = VIBER::choose_clusters(Set7_standard$fit_VIBER)

pdf("VIBER_Set7_fits_heuristics.pdf", width = 4, height = 3)
lapply(VIBER:::plot.vb_bmm(reduced_mv), print)
dev.off()
```

