---
title: "Query, filter, re-shape and plot datasets"
author: "Giulio Caravagna"
date: "September 2019"
institute: "Institute for Cancer Research"
email: "giulio.caravagna@icr.ac.uk"
output: rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Query, filter, re-shape and plot datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows query, filter, re-shape and plot operations on the data stored inside a dataset of the mvMOBSTER package (S3 object class `mbs_data`).

We work with the `x` generated in the introductory vignette "Getting started". 

```{r, include=FALSE, echo=FALSE, message= FALSE, warning=FALSE}
library(mvmobster)

# Simple fake dataset
data('example_input_formats', package = 'mvmobster')

x = dataset(
  mutations = example_input_formats$long_mutations, 
  segments = example_input_formats$long_segments, 
  samples = c('Set42', 'Set44'), 
  purity =  c(`Set42` = 0.8, `Set44` = 0.8), 
  description = "A fake dataset"
)
```

### Query the data

Getters functions take in input `x` and usually the list of samples to query (all by default), and ids to subset the call to a specific set of mutations.

```{r}
# Number of mutations in the data, and number of samples
N(x)
S(x)

# VAF of the data
VAF(x)

# VAF of one specific sample
VAF(x, samples = x$samples[1])

# Same access for read counts
DP(x)
NV(x)
```

Getters can be easily assembled joining them by id, as done by function

```{r}
print(Data_table)

# All data table, with annotations as well
Data_table(x)
```

### Plotting the data

The basic S3 `plot` function returns a list for every pair of input samples. If there is one sample, a histogram is plotted.
```{r, fig.width=3, fig.height=3}
plot(x)
```

A report function visualises all the data available for a sample.
```{r, fig.width=8, fig.height=9}
plot_sample(x, x$samples[1])
```

The CNAqc objects generated when data have been loaded, are available inside `x`, and can be processed wih CNAqc functions.

```{r, fig.width=3, fig.height=4}
# CNAqc: plot percentage of genome at each karyotype.
plot_karyotypes(x$CNAqc$Set42)
```



