---
title: "Cancer Evolution Analysis"
author: "Giulio Caravagna"
date: "September 2019"
institute: "Institute for Cancer Research"
email: "giulio.caravagna@icr.ac.uk"
output: rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Cancer Evolution Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows the analysis with MOBSTER and VIBER of the data stored inside a dataset of the mvMOBSTER package (S3 object class `mbs_data`).

We work with the `x` generated in the introductory vignette "Getting started". 

```{r, include=FALSE, echo=FALSE, message= FALSE, warning=FALSE}
library(mvmobster)

# Simple fake dataset
data('example_input_formats', package = 'mvmobster')

x = dataset(
  mutations = example_input_formats$long_mutations, 
  segments = example_input_formats$long_segments, 
  samples = c('Set42', 'Set44'), 
  purity =  c(`Set42` = 0.8, `Set44` = 0.8), 
  description = "A fake dataset"
)
```

### MOBSTER analysis

The wrapper `analyze_mobster` runs `mobster_fit` from the [MOBSTER](https://caravagn.github.io/MOBSTER/) package for tumour subclonal deconvolution, which combines Dirichlet  mixtures with distributions predicted by theoretical population genetics.

If you want to pass extra parameters to the fitting function, please refer to  [MOBSTER vignettes ](https://caravagn.github.io/MOBSTER/) on possible parametrizations of the fit.

```{r fig.width=8, fig.height=9}
x = analyze_mobster(x, parallel = F, K = 1,  init = 'random')
```

Now the internal state of the object shows the fits as available analysis.
```{r fig.width=8, fig.height=9}
print(x)
```

Consequently, the S3 plotting function can now be asked to plot 2D points coloured by MOBSTER cluster (the samples are analyzed independently therefore there are two labels per point).

```{r fig.width=3, fig.height=3}
plot(x, clusters = 'MOBSTER')
```

The MOBSTER fits are accessible
```{r fig.width=4, fig.height=4}
plot(x$fit_MOBSTER$Set42$best)
plot(x$fit_MOBSTER$Set44$best)
```



x_notail = filter_tails(x)
x_notail = analyze_VIBER(x_notail)

plot(x_notail, clusters = 'VIBER')

```{r fig.width=4, fig.height=4}
x = analyze_VIBER(x)
print(x)
plot(x, clusters = 'VIBER')
```
